\subsection{Additional Heuristics}
\label{sec:heuristics}

Beyond the primary algorithm categories, three additional heuristics from the
MWC literature are implemented that offer different trade-offs between solution
quality and computational efficiency.

\subsubsection{Semi-Exact Heuristic with Graph Reduction}
\label{sec:fast-wclq}

FastWClq~\cite{fang2021fastwclq} combines randomized clique construction with
graph reduction. Its key innovation is the Best from Multiple Selection (BMS)
strategy, which samples $k$ candidates and selects the best one based on a
benefit function.

\begin{algorithm}[H]
    \caption{FastWClq}
    \label{alg:fast-wclq}
    \footnotesize
    \scalebox{0.65}{
        \begin{minipage}{0.85\linewidth}
            \begin{algorithmic}[1]
                \Require Graph $G$, BMS parameter $k$, time\_limit
                \Ensure Best clique $C^*$ found
                \State $G' \gets G$, $C^* \gets \emptyset$, $W^* \gets 0$
                \While{$V(G') \neq \emptyset$ and not timeout}
                \State $C \gets$ \textsc{BMSConstruction}($G'$, $k$)
                \If{$W(C) > W^*$}
                \State $C^* \gets C$, $W^* \gets W(C)$
                \ForAll{$v \in V(G')$} \textbf{if} $W(N[v]) \leq W^*$ \textbf{then} remove $v$ \EndFor
                \EndIf
                \EndWhile
                \State \Return $C^*$
            \end{algorithmic}
        \end{minipage}
    }
\end{algorithm}

The BMS construction selects vertices based on: $\text{benefit}(v) = w(v) + 0.1
    \cdot \sum_{u \in N(v) \cap candidates} w(u)$

\textbf{Complexity:} $O(T \cdot n \cdot k)$ per iteration, where $T$ is iterations until graph becomes empty.

\subsubsection{Local Search with Strong Configuration Checking}
\label{sec:scc-walk}

SCCWalk~\cite{wang2020sccwalk} uses local search with Strong Configuration
Checking (SCC) to avoid cycling and walk perturbation to escape local optima.

\begin{algorithm}[H]
    \caption{SCCWalk}
    \label{alg:scc-walk}
    \footnotesize
    \scalebox{0.65}{
        \begin{minipage}{0.85\linewidth}
            \begin{algorithmic}[1]
                \Require Graph $G$, max\_unimprove $M$, perturbation strength $\sigma$
                \Ensure Best clique $C^*$ found
                \State $C \gets$ \textsc{GreedyConstruct}($G$), $C^* \gets C$, $steps \gets 0$
                \While{not stopping criteria}
                \State Try Add, Swap, or Drop operations using SCC rules
                \If{$W(C) > W(C^*)$} $C^* \gets C$, $steps \gets 0$
                \Else~$steps \gets steps + 1$
                \EndIf
                \If{$steps \geq M$} $C \gets$ \textsc{WalkPerturbation}($C$, $\sigma$), $steps \gets 0$ \EndIf
                \EndWhile
                \State \Return $C^*$
            \end{algorithmic}
        \end{minipage}
    }
\end{algorithm}

SCC prevents revisiting configurations by tracking vertex addition/removal
timestamps.

\textbf{Complexity:} $O(T \cdot n^2)$ for $T$ iterations, each involving neighborhood exploration.

\subsubsection{Hybrid Reduction with Peeling}
\label{sec:mwc-peel}

MWCPeel combines exact reduction rules with heuristic ``peeling'' that removes
low-score vertices, progressively simplifying the graph.

\begin{algorithm}[H]
    \caption{MWCPeel}
    \label{alg:mwc-peel}
    \footnotesize
    \scalebox{0.65}{
        \begin{minipage}{0.85\linewidth}
            \begin{algorithmic}[1]
                \Require Graph $G$, peel\_fraction $\phi$
                \Ensure Best clique $C^*$ found
                \State $G' \gets G$, $C^* \gets$ \textsc{InitialClique}($G$)
                \While{$V(G') \neq \emptyset$}
                \State Apply exact reduction rules (domination, etc.)
                \State Compute scores: $score(v) = W(N[v])$
                \State Peel bottom $\phi$ fraction of vertices by score
                \EndWhile
                \State Solve remaining small graph exactly
                \State \Return $C^*$
            \end{algorithmic}
        \end{minipage}
    }
\end{algorithm}

\textbf{Complexity:} $O(n^3)$ for reductions plus $O(2^{n'})$ for final exact solve on small graph.

\subsubsection{Summary of Heuristics}

The following table summarizes these complementary heuristics:

\begin{table}[H]
    \centering
    \caption{Summary of Additional Heuristics}
    \label{tab:heuristics-summary}
    \begin{tabular}{lccc}
        \toprule
        \textbf{Algorithm} & \textbf{Type} & \textbf{Strength}    & \textbf{Weakness}  \\
        \midrule
        FastWClq           & Semi-exact    & Can prove optimality & May not converge   \\
        SCCWalk            & Local search  & Good local optima    & Stuck in basins    \\
        MWCPeel            & Reduction     & Fast preprocessing   & Aggressive peeling \\
        \bottomrule
    \end{tabular}
\end{table}
