\section{Introduction}
\label{sec:introduction}

The Maximum Weight Clique (MWC) problem is a fundamental combinatorial
optimization problem with applications spanning social network analysis,
bioinformatics, computer vision, and resource
scheduling~\cite{bomze1999maximum}. Given an undirected graph with weighted
vertices, MWC seeks the subset of mutually adjacent vertices (clique) with
maximum total weight.

As an NP-hard problem~\cite{garey1979computers}, MWC admits no polynomial-time
exact algorithm unless P = NP. This intractability has motivated extensive
research into both exact algorithms with exponential worst-case complexity and
heuristic/approximation methods that trade optimality for
efficiency~\cite{wu2015review}.

This work presents a comprehensive study of algorithms for MWC, with particular
emphasis on \textbf{randomized approaches}. Randomized
algorithms~\cite{motwani1995randomized} use random choices during execution,
offering probabilistic guarantees or empirically good performance. Two
paradigms are distinguished:
\begin{itemize}
      \item \textbf{Monte Carlo algorithms:} Run in polynomial time but may return incorrect results with bounded probability.
      \item \textbf{Las Vegas algorithms:} Always return correct results but with variable (potentially unbounded) runtime.
\end{itemize}

The contributions of this work include:
\begin{enumerate}
      \item Implementation and analysis of 14 algorithms spanning four categories:
            \begin{itemize}\sloppy
                  \item Randomized algorithms (random construction, hybrid approaches, iterative
                        search, Monte Carlo, Las Vegas)
                  \item Reduction-based methods (MWCRedu, \\ MaxCliqueWeight variants)
                  \item Exact branch-and-bound (WLMC, TSM-MWC)
                  \item Additional heuristics (FastWClq, SCCWalk,\\ MWCPeel)
            \end{itemize}
      \item Comprehensive experimental evaluation on generated graphs of varying sizes and
            densities
      \item Analysis of solution quality, execution time, and scalability trade-offs
      \item Practical recommendations for algorithm selection based on problem
            characteristics
\end{enumerate}

The remainder of this paper is organized as follows: Section~\ref{sec:problem}
formally defines the MWC problem. Section~\ref{sec:algorithms} presents the
implemented algorithms with pseudocode and complexity analysis.
Section~\ref{sec:results} describes the experimental methodology and results.
Section~\ref{sec:discussion} analyzes trade-offs and failure modes.
Section~\ref{sec:conclusion} summarizes findings and suggests future work.
