\begin{thebibliography}{00}

    \bibitem{cormen2009} T. H. Cormen, C. E. Leiserson, R. L. Rivest, and C. Stein, \textit{Introduction to Algorithms}, 3rd ed.
    MIT Press, 2009.
    
    \bibitem{garey1979} M. R. Garey and D. S. Johnson, \textit{Computers and Intractability: A Guide to the Theory of NP-Completeness}.
    W. H. Freeman, 1979.
    
    \bibitem{erhardt2023improved} R. Erhardt, K. Hanauer, N. M. Kriege, C. Schulz, e D. Strash, "Improved Exact and Heuristic Algorithms for Maximum Weight Clique," \textit{arXiv preprint arXiv:2302.00458}, Fev. 2023. [6, 9, 14, 15, 16]
    
    \bibitem{konc2025efficient} J. Konc e D. Janežič, "Efficient Algorithms for the Maximum Weight Clique Problem with Applications to Protein Binding Site Analysis," in \textit{Proceedings of the 13th MATCOS Conference (MATCOS 2025)}, 2025 (previsto). [4]
    
    \bibitem{sun2021ml} Y. Sun, X. Li, e A. Ernst, "Using Statistical Measures and Machine Learning for Graph Reduction to Solve Maximum Weight Clique Problems," \textit{IEEE Transactions on Pattern Analysis and Machine Intelligence}, vol. 43, no. 5, pp. 1746–1760, Mai. 2021. [8]
    
    \bibitem{sun2024adaptive} Y. Sun, S. Esler, D. Thiruvady, A. T. Ernst, X. Li, e K. Morgan, "Adaptive Population-based Simulated Annealing for Resource Constrained Job Scheduling with Uncertainty," \textit{International Journal of Production Research}, vol. 62, no. 17, pp. 6227–6250, 2024. [8]
    
    \bibitem{rcpred2019} A. Legendre, E. Angel, e F. Tahi, "RCPred: RNA complex prediction as a constrained maximum weight clique problem," \textit{BMC Bioinformatics}, vol. 20-S(3), pp. 53-62, 2019. (Citado como ferramenta ativa em relatórios de 2022-2023 \cite{apogeebio2023}).
    
    \bibitem{apogeebio2023} ApogeeBIO Research Team, "Hosting Teams Report 2021-2023," \textit{ApogeeBIO}, Dez. 2023. [2, 5]
    
    \bibitem{hao2023metaheuristic} Y. Wang e J. Hao, "Metaheuristic algorithms," in \textit{Discrete Diversity and Dispersion Maximization - A Tutorial on Metaheuristic Optimization}, R. Martí, A. Martínez-Gavara, Eds. Springer, 2023, pp. 271-298. [7]
    
\end{thebibliography}